---
title: "quiz 2"
output: html_document
date: "2025-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(readr)
library(ggplot2)
library(ggfortify)
library(MASS)
```

```{r}
Glakes <- read_csv("C:/mat374/Glakes.csv")
head(Glakes)
```

a,
```{r}
model1 <- lm(Time ~ Tonnage, data = Glakes)
summary(model1)

autoplot(model1, label.repel = TRUE, label.size = 3)

```
```{r}
predict(model1, newdata = data.frame(Tonnage = 10000), interval = "confidence", level = 0.95)
```

The first model is Time = 12.3447 + 0.0065 * Tonnage + ε.
The R-squared is 0.839, indicating 83.9% of the variability in Time is explained by Tonnage.
The predicted time when tonnage is 10,000 is 77.52, with a 95% confidence interval of [69.36, 85.68].

b,
```{r}
predict(model1, newdata = data.frame(Tonnage = 160000)) - predict(model1, newdata = data.frame(Tonnage = 10000))
```

The predicted change in time when tonnage increases from 10,000 to 160,000 is 977.68.

c,
```{r}
model2 <- lm(I(log(Time)) ~ I(Tonnage^0.25), data = Glakes)
summary(model2)

autoplot(model2, label.repel = TRUE, label.size = 3)

```
```{r}
pred_log <- predict(model2, newdata = data.frame(Tonnage = 10000), interval = "confidence", level = 0.95)
exp(pred_log)
```
The second model is log(Time) = 1.19 + 0.31 * Tonnage^{0.25}.
The R-squared is 0.816.
The predicted time when tonnage is 10,000 is 72.20, with a 95% confidence interval of [58.64, 88.89].

d,
```{r}
pred_log_10000 <- predict(model2, newdata = data.frame(Tonnage = 10000))
pred_log_160000 <- predict(model2, newdata = data.frame(Tonnage = 160000))
exp(pred_log_160000) - exp(pred_log_10000)
```

```{r}
coef(model2)
exp(coef(model2) * 10000^0.25)
```

The predicted change in time when tonnage increases from 10,000 to 160,000 is 1516.03.
For a 16-fold increase in tonnage (from x to 16x), the ratio y2 / y1 = exp(β₁ * x^{0.25}) = exp(0.31 * x^{0.25}).
For x=10,000, the ratio is approximately 22.

e,
```{r}
boxcox(model1)
```

Yes, the second model is an improvement over the first in terms of predicting Time.
The Box-Cox transformation suggests λ = 0.5, indicating a log transformation on Time is appropriate, aligning with the second model.