---
title: "quiz3"
output: html_document
date: "2025-10-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r cars}
library(readr)
library(ggplot2)
library(ggfortify)
library(MASS)
Copper <- read_csv("C:/mat374/Copper.csv")
```

1. 
```{r}
ggplot(data = Copper, aes(x = Sulfur, y = Tension)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

copper_lm <- lm(Tension ~ Sulfur, data = Copper)
summary(copper_lm)

autoplot(copper_lm, label.repel = TRUE, label.size = 3)
```

Residuals vs Fitted: Clear curvature suggests the linearity assumption is violated.
Scale-Location: Show increasing spread with fitted values.
Normal Q-Q: deviations at the tails, indicating the errors is not normally distributed.
Residuals vs Leverage: No extreme outliers, but some high leverage points are present.
Linearity, equal variances, and normality are violated

2.
```{r}
copper_lm_sqrt <- lm(Tension ~ I(sqrt(Sulfur)), data = Copper)
summary(copper_lm_sqrt)

autoplot(copper_lm_sqrt, label.repel = TRUE, label.size = 3)
```

The estimated regression equation is: Tension = 123.17 + 742.35 * sqrt(Sulfur) + e.

3.
```{r}
pred_01 <- predict(copper_lm_sqrt, newdata = data.frame(Sulfur = 0.1))
pred_02 <- predict(copper_lm_sqrt, newdata = data.frame(Sulfur = 0.2))
pred_02 - pred_01
```

